Temporal stability in human interaction networks: sector sizes, topological prominence and time activity

### 2015-10-04 paper.tex
revisado artigo ateh as conclusoes

The covariance matrix 
$\mathbf{C}$ is the correlation matrix because $\mathbf{X'}$ is normalized.
Therefore, $\mathbf{C}$ is also directly observed as a first clue for patterns
by the most simple associations:
low absolute values indicate low correlation (and a possible independence);
high values indicate positive correlation;
negative values with a high absolute value indicate negative correlation.

@@

\noindent where $\eta_k$ is the number of vertices with degree $k$, while $\overline{\Delta}_{i-1}=max(\Delta_{i-1})$, and $\overline{\Delta}_{0}=-1$. Equation~\ref{criterio} can now be written in the form:

\begin{equation}\label{criterio2}
	\begin{split}
	\sum_{x=min(\Delta_i)}^{max(\Delta_i)} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{max(\Delta_i)} P(x) \Rightarrow \\
	\Rightarrow \Delta_i \text{ holds intermediary degree values.}
	\end{split}
\end{equation}

@@

\subsection{Evolution and visualization of the networks}\label{sec:viz}
The evolution of the networks was observed within sequences of snapshots. with a fixed number of messages,
Each of these sequences had a fixed number of messages, the window size $ws$,
which was considered with different shifts in the message timeline to obtain snapshots.
In each snapshot was performed both the PCA with topological metrics and the Erd\"os sectioning.  
The values of $ws$ employed were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000.
Variations in the number of vertices, edges
and other network characteristics
are exposed in Section~\ref{si:frac}
of the Supporting Information document.

@@

\begin{table}
	\caption{Activity percentages of the months along the year from LAU list. Activity is usually concentrated in Jun-Aug and/or in Dec-Mar (see Section~\ref{si:months} of the Supporting Information). These observations fit academic calendars, vacations and end-of-year holidays.}
	\footnotesize
	\input{tables/tabMonthsLAU}
	\label{tab:min2}
\end{table}
\subsection{Stability of principal components and the prevalence of symmetry over clusterization for dispersion}\label{prevalence}
%The topology was analyzed using standard, well-established metrics of centrality and clustering.
%We also introduced symmetry metrics given the evidence of their importance in social contexts~\cite{newmanEvolving}.
The contribution of each metric to the variance is very similar for all the networks and along time.
In applying PCA to the snapshots, the contribution of each metric to the principal components presents very small standard deviation.
Table~\ref{tab:pcain} exemplifies the formation of principal components for the MET email list.
Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
of the Supporting Information document for the other email lists (benchmarks) and other interaction networks,
with the consideration of strategic combinations of metrics.

@@

\subsection{Scalable fat-tail structure: constancy of membership fractions in each Erd\"os sector}\label{subsec:pih}

\begin{figure*} 
	\centering
	\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
	\caption{Fractions of agents in each Erd\"os sector, where the fractions for hubs, intermediary and peripheral vertices are represented in red, green and blue, respectively. We used two simple criteria, namely degree and strength, for the graphics on the left. For the graphics on the right we employed the Exclusivist and Inclusivist compound criteria, with black lines representing the fraction of vertices without class and with more than one class, respectively. See Section~\ref{si:frac} of Supporting Information for a collection of such timeline figures with all simple and compound criteria and metrics. Table~\ref{tab:secE}, also from Supporting Information, presents these fractions of agents in snapshots of networks from Facebook, Twitter and Participabr.}
	\label{fig:sectIL}
\end{figure*}

\subsection{Types from Erd\"os sectors}\label{sec:pty}
A sector to which a vertex belongs can be regarded as yielding a type to the corresponding participant.
Assigning a type to a participant inevitably raises an important question regarding the possible stigmatization. 
We take the view that the participation typology inherent in the Erd\"os sectors is not stigmatizing because the type of an individual changes along time and as different networks are considered~\cite{adorno}. 
That is to say, an individual is a hub in a number of networks and peripheral in other networks, and even within a network he/she probably changes type along time. 
Indeed, we did observe often transitions of participants from one sector to another.
The typology proposed here bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

We analyzed the time evolution of the networks using visualization tools developed for this research~\cite{rcText,versinus} and inspected the raw data to infer the main characteristics of each type. Our main observations may be summarized as follows:
\begin{itemize}
	\item Core hubs usually have intermittent activity. Very stable activity was found on MET hubs, which is consistent with the literature where greater stability occurs in smaller communities~\cite{barabasiEvo}.
	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone. The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions. Therefore, intermediary vertices seem responsible for the network structure. Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs. These seem as authorities and contrast with participants that respond much more than receive responses.
	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members known each other often, is found mostly in peripheral and intermediary sectors.
\end{itemize}

With regard to the networks as the whole objects of analysis,
we were able to observe a negative correlation between the number of threads and the number of participants.
When the number of participants exceeds a threshold,
the number of threads displays a positive correlation with the number of participants.
This finding is illustrated in Figure~\ref{fig:nmgamma3d} and can also be observed in Table~\ref{tab:genLists}.
Obviously, network types can be derived from such results, which was not attempted here but 
left for the reader and future work.

\begin{figure}
   \centering
        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists. Highest number of threads are found in lists with few participants. The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise. This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}. For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants. All LAU, LAD, MET lists present smaller networks.}
	\label{fig:nmgamma3d}
\end{figure}

>>>>

@@

\noindent where $\eta_k$ is the number of vertices with degree $k$, while $\overline{\Delta}_{i}=max(\Delta_{i})$, and $\overline{\Delta}_{0}=-1$. Equation~\ref{criterio} can now be written in the form:

\begin{equation}\label{criterio2}
	\begin{split}
		\sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} \widetilde{P}(x) < \sum_{x=min(\Delta_i)}^{\overline{\Delta}_i} P(x) \Rightarrow \\
	\Rightarrow \Delta_i \text{ holds intermediary degree values.}
	\end{split}
\end{equation}

@@

The covariance matrix 
$\mathbf{C}$ is the correlation matrix because $\mathbf{X'}$ is normalized.
Therefore, $\mathbf{C}$ is also directly observed as a first clue for patterns
by the most simple associations:
low absolute values indicate low correlation (and a possible independence);
high values indicate positive correlation;
negative values with a high absolute value indicate negative correlation.
\subsection{Evolution and visualization of the networks}\label{sec:viz}
The evolution of the networks was observed within sequences of snapshots. with a fixed number of messages,
Each of these sequences had a fixed number of messages, the window size $ws$,
which was considered with different shifts in the message timeline to obtain snapshots.
In each snapshot was performed both the PCA with topological metrics and the Erd\"os sectioning.  
The values of $ws$ employed were 50, 100, 200, 400, 500, 800, 1000, 2000, 2500, 5000 and 10000.
Variations in the number of vertices, edges
and other network characteristics
are exposed in Section~\ref{si:frac}
of the Supporting Information document.

@@

\begin{table}
	\caption{Activity percentages of the months along the year from LAU list. Activity is usually concentrated in Jun-Aug and/or in Dec-Mar (see Section~\ref{si:months} of the Supporting Information). These observations fit academic calendars, vacations and end-of-year holidays.}
	\footnotesize
	\input{tables/tabMonthsLAU}
	\label{tab:min2}
\end{table}

\subsection{Stable fractions of participants in each Erd\"os sector}\label{subsec:pih}

\begin{figure*} 
	\centering
	\includegraphics[width=\textwidth]{figs/InText-WLAU-S1000}
	\caption{Fractions of agents in each Erd\"os sector, where the fractions for hubs, intermediary and peripheral vertices are represented in red, green and blue, respectively. Fractions derived from degree and strength criteria are both on the left.
	Fractions derived from the exclusivist $C_1$ and the inclusivist $C_2$ compound criteria are exposed in the plots to the right,
with black lines representing the fraction of vertices without class and with more than one class, respectively. 
See Section~\ref{si:frac} of Supporting Information document for a collection of such timeline figures with all simple and compound criteria specified in Section~\ref{sectioning}.
Table~\ref{tab:secE}, also from Supporting Information document, presents these fractions of agents in networks from Facebook, Twitter and Participabr.}
	\label{fig:sectIL}
\end{figure*}

\subsection{Stability of principal components and the prevalence of symmetry over clusterization for dispersion}\label{prevalence}
%The topology was analyzed using standard, well-established metrics of centrality and clustering.
%We also introduced symmetry metrics given the evidence of their importance in social contexts~\cite{newmanEvolving}.
The contribution of each metric to the variance is very similar for all the networks and along time.
In applying PCA to the snapshots, the contribution of each metric to the principal components presents very small standard deviation.
Table~\ref{tab:pcain} exemplifies the formation of principal components for the MET email list.
Similar results are presented in Sections~\ref{si:pcat} and~\ref{si:ext}
of the Supporting Information document for the other email lists (benchmarks) and other interaction networks,
with the consideration of strategic combinations of metrics.

@@

\subsection{Types from Erd\"os sectors}\label{sec:pty}
The type of the Erd\"os sector to which a participant belongs 
can be regarded as implying a type to the corresponding participant.
Assigning a type to a participant raises important 
issues about the potential for stigmatization and prejudice. 
In the present case, the type of a participant
changes both along time and as
different networks are considered,
and therefore the potential for prejudice of such
participant typology is attenuated~\cite{adorno}. 
In other words, an individual is a hub in a number of networks and peripheral in other networks, and even within the same network he/she probably changes type along time~\cite{4hubs}.
Moreover, such proposed typology of participants bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

We analyzed the temporal evolution of the networks using visualization tools developed for this research~\cite{rcText,versinus} and inspected the raw data to infer the main characteristics of each type. Our main observations may be summarized as follows:
\begin{itemize}
	\item Core hubs usually have intermittent activity. Very stable activity was found only in smaller communities, which is consistent with the literature~\cite{barabasiEvo}.
	\item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone.
The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions.
Therefore, intermediary vertices seem responsible for the network structure.
Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity, to point just a few examples of this greater variety of roles.
	\item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs.
These seem as authorities and contrast with participants that respond much more than receive responses.
	\item The most obvious community structure, as observed by a high clustering coefficient, i.e. members know each other often, is found mostly in peripheral and intermediary sectors.
\end{itemize}

With focus on the networks as the whole objects of analysis,
we were able to observe a peculiar correlation pattern 
between the number of threads and the number of participants.
When the number of participants exceeds a threshold,
the number of threads displays a positive correlation with the
number of participants.
This finding is illustrated in Figure~\ref{fig:nmgamma3d} and can also be observed in Table~\ref{tab:genLists}.
Obviously, network types can be derived from such results, which was not attempted here but 
left for future work.

\begin{figure}
   \centering
        \includegraphics[width=\columnwidth]{figs/mpgamma____________}
	\caption{A scatter plot of number of messages (M) versus number of participants (N) versus number of threads ($\Gamma$) for 140 email lists. Highest numbers of threads are found in lists with few participants. The correlation between $N$ and $\Gamma$ is negative for low values of $N$ but positive otherwise. This negative correlation between $N$ and $\Gamma$ can also be observed in Table~\ref{tab:genLists}. For $M=20000$ messages, positive correlation of $N$ and $\Gamma$ is present mostly above 1500 participants. All LAU, LAD, MET lists present smaller networks.}
	\label{fig:nmgamma3d}
\end{figure}
